---
title: "Problem set grade summaries"
format:
  html:
    toc: false
    number-sections: false
editor: source
editor_options:
  chunk_output_type: console
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

library(tidyverse)

source("src/r/grading_pipeline.R")

# Everything below is read from the saved grading run, so re-rendering after
# more marking is finished brings the numbers with it.

run <- read_grading_results(1)

students <- run$students

questions <- run$questions

slots <- run$slots

question_text <-
  questions_from_qmd(1) %>%
  transmute(
    question,
    stem = str_squish(str_remove_all(stem, "[*`]"))
  )

# A part's own wording, so a table can say what was being asked rather than
# only which number it was.

part_text <-
  slots %>%
  distinct(
    question,
    answer_order,
    criterion
  ) %>%
  left_join(question_text, by = "question") %>%
  mutate(
    asked = str_squish(str_remove_all(coalesce(criterion, stem), "[*`]"))
  )

standard_table <-
  function(.data) {
    .data %>%
      kableExtra::kbl() %>%
      kableExtra::kable_paper("hover", full_width = FALSE)
  }
```

<button class="accordion">Problem set 1</button>

::: panel

:::{style="background-color: white; font-size: 14px;"}

## Problem set 1: R foundations

```{r}
#| label: provisional
#| echo: false
#| message: false
#| warning: false
#| results: asis

flagged <- sum(slots$needs_review)

cat(
  "::: callout-note\n",
  "These numbers are provisional. ", flagged,
  " answers are still being looked at by hand, and every one of them is",
  " currently scored as **correct**, so a few marks may still come down.\n",
  ":::\n",
  sep = ""
)
```

Descriptive statistics of scores for Problem set 1:

```{r}
#| label: descriptives
#| echo: false
#| message: false
#| warning: false
#| results: asis

cat(
  "* Submissions: ", nrow(students), "\n",
  "* Mean: ", round(mean(students$total), 2), "\n",
  "* Median: ", round(median(students$total), 2), "\n",
  "* Standard deviation: ", round(sd(students$total), 2), "\n",
  "* Standard error: ",
  round(sd(students$total) / sqrt(nrow(students)), 2), "\n",
  "* Range: ", round(min(students$total), 2), " to ",
  round(max(students$total), 2), "\n",
  sep = ""
)
```

Scores are out of 10, plus up to 1 point of extra credit for following the
course style guide.

Here is the distribution of scores:

```{r}
#| label: distribution
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4

students %>%
  ggplot() +
  aes(x = total) +
  geom_histogram(
    binwidth = 0.5,
    fill = "#dcdcdc",
    color = "black"
  ) +
  geom_vline(
    xintercept = mean(students$total),
    linetype = "dashed"
  ) +
  scale_x_continuous(
    breaks = seq(0, 11, 1),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  coord_cartesian(xlim = c(0, 11.1)) +
  labs(
    x = "Score (out of 10, plus style credit)",
    y = "Students"
  ) +
  theme_classic()
```

The dashed line is the class mean.

### Which questions were hardest

```{r}
#| label: by_question_plot
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.45

by_question <-
  questions %>%
  group_by(question) %>%
  summarize(
    points = first(question_points),
    share_lost = mean(deduction / question_points),
    .groups = "drop"
  )

by_question %>%
  ggplot() +
  aes(
    x = factor(question),
    y = share_lost
  ) +
  geom_col(
    fill = "#dcdcdc",
    color = "black"
  ) +
  scale_y_continuous(
    labels = scales::percent,
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Question",
    y = "Average share of the question's points lost"
  ) +
  theme_classic()
```

```{r}
#| label: by_question_table
#| echo: false
#| message: false
#| warning: false

by_question %>%
  left_join(question_text, by = "question") %>%
  arrange(desc(share_lost)) %>%
  transmute(
    Question = question,
    Points = points,
    `Average deduction (%)` = round(share_lost * 100, 1),
    `What it asked` = str_trunc(stem, 90)
  ) %>%
  standard_table()
```

### The parts that gave the most trouble

Several questions were made of a few separate parts. These are the individual
parts that cost the class the most.

```{r}
#| label: hardest_parts
#| echo: false
#| message: false
#| warning: false

slots %>%
  group_by(question, answer_order) %>%
  summarize(
    points = first(slot_points),
    share_lost = mean(deduction / slot_points),
    .groups = "drop"
  ) %>%
  left_join(
    part_text,
    by = c("question", "answer_order")
  ) %>%
  filter(share_lost > 0) %>%
  arrange(desc(share_lost)) %>%
  head(8) %>%
  transmute(
    Part = str_c(question, ".", answer_order),
    Points = points,
    `Average deduction (%)` = round(share_lost * 100, 1),
    `What it asked` = str_trunc(asked, 90)
  ) %>%
  standard_table()
```

### Using the functions the assignment allows

This was the single biggest source of lost points, and it runs through most of
the questions below. Each problem set lists the functions you may use, and
reaching outside that list costs half the question, however good the answer is
otherwise.

The extraction operator `$` accounted for most of it. The hints in the problem
set called it out by name, and it still came up more than any other function.

```{r}
#| label: unapproved
#| echo: false
#| message: false
#| warning: false

used_outside <-
  slots %>%
  select(student_id, unapproved) %>%
  unnest(unapproved)

if (nrow(used_outside)) {
  used_outside %>%
    group_by(unapproved) %>%
    summarize(
      Uses = n(),
      Students = n_distinct(student_id),
      .groups = "drop"
    ) %>%
    arrange(desc(Students), desc(Uses)) %>%
    rename(Function = unapproved) %>%
    standard_table()
}
```

`[[` does everything `$` does for these questions, and it is on the list.
`filter()` is a fine way to subset rows in general; it just was not among the
functions available for this problem set, which is about the operators
underneath the tidyverse rather than the tidyverse itself.

A useful habit: before submitting, read your answer against the list and check
every function in it, operators included.

### The reference tree

Question 4 asked for a reference tree showing the objects associated with names
in your global environment. That is the environment itself, `.GlobalEnv`, not
one object inside it.

```{r}
#| label: reference_tree
#| echo: false
#| message: false
#| warning: false

slots %>%
  filter(question == 4, answer_order == 4) %>%
  mutate(
    target =
      if_else(
        str_detect(answer, "GlobalEnv"),
        "lobstr::ref(.GlobalEnv)",
        "pointed at something else"
      )
  ) %>%
  count(target, name = "Students") %>%
  rename(`What was passed to ref()` = target) %>%
  standard_table()
```

### The style guide

Style is extra credit: 10% of a question's points, added when every answer in
that question follows the guide.

```{r}
#| label: style_credit
#| echo: false
#| message: false
#| warning: false
#| results: asis

cat(
  "* Average style credit earned: ",
  round(mean(students$style_credit), 2), " of 1.00\n",
  "* Students earning all of it: ",
  sum(near(students$style_credit, 1)), " of ", nrow(students), "\n",
  sep = ""
)
```

These are the conventions the class broke most often:

```{r}
#| label: style_violations
#| echo: false
#| message: false
#| warning: false

slots %>%
  select(student_id, violations) %>%
  unnest(violations) %>%
  group_by(comment_text) %>%
  summarize(
    Times = n(),
    Students = n_distinct(student_id),
    .groups = "drop"
  ) %>%
  arrange(desc(Times)) %>%
  head(8) %>%
  rename(Convention = comment_text) %>%
  standard_table()
```

Question 10 handed you a block with ten specific violations in it and asked you
to repair them. These are the ones most often left in place:

```{r}
#| label: repairs_missed
#| echo: false
#| message: false
#| warning: false

repairs <-
  slots %>%
  filter(question == 10) %>%
  filter(!is.na(repair_unfixed_names), repair_unfixed_names != "")

if (nrow(repairs)) {
  tibble(
    fault = unlist(str_split(repairs$repair_unfixed_names, " \\| "))
  ) %>%
    count(fault, name = "Students") %>%
    arrange(desc(Students)) %>%
    rename(`Violation left unrepaired` = fault) %>%
    standard_table()
}
```

```{r}
#| label: repairs_clean
#| echo: false
#| message: false
#| warning: false
#| results: asis

clean <- sum(slots$question == 10 & slots$repair_unfixed == 0)

cat(
  clean, " of ", nrow(students),
  " submissions repaired all ten.\n",
  sep = ""
)
```

:::

:::
